Time series data are ubiquitous across many different fields.
% Similarity measures are the core components.
Many data mining tasks, such as classification, clustering, and motif finding, have been defined for time series data. 
They utilize similarity search as a core subroutine, making it crucial to design similarity measures that align with our intuitions.
%
To facilitate efficient computation, speedup techniques are essential.
%
% DTW
Dynamic Time Warping (DTW) is arguably the most prevailing distance measure for time series data. 
% US
However, studies have shown that for certain data, another distance measure, namely Uniform Scaling (US), is equally crucial as DTW.
DTW handles the local distortion, while US handles the global scaling.
% DTW + US
In addition, studies have demonstrated that combining DTW and US is necessary to obtain meaningful results in some cases. 
%
% Motivation
Surprisingly, all existing studies employ only a single scaling factor for the entire time series.
A time series could consist of phases. 
Since each phase of a time series expresses at its own rate, using a single scaling factor is insufficient when comparing two time series that share similar phases but differ in their expression rates.
%
% Contribution
We introduce the first framework that accounts for multiple scaling factors, Piecewise Scaling Distance (PSD).
PSD employs other existing distance measures as subroutines.
Because the naive implementation of PSD is slow, we propose a constrained version of PSD that enforces constraints based on the allowed segment lengths derived from the given scaling factor bound.
It also prevents pathological results.
In addition, two other speedup techniques have been proposed, which achieve 10.10X to 191.46X speedup.
We also demonstrate the usage of a lower bound when DTW is used as the subroutine of PSD.
% It can also help us better understand the data by dividing the long time series into several homogeneous segments.
Moreover, we show that the segmentation results returned by PSD can improve the accuracy of other distance measures.
% Code is available at: \url{https://github.com/cyuab/k-scaling-factor-dtw}.